---
title: "Introduction to OregonFrogs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{oregonfrogs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  out.width = "100%",
  warning = FALSE,
  message = FALSE,
  fig.align = 'center'
)
```

The `oregonfrogs` R package contains two datasets:

- `oregonfrogs_raw`
- `oregonfrogs`

In addition, it provides two useful functions for spatial data manipulation:

- `longlat_to_utm()`
- `utm_to_longlat()`


```{r setup}
library(oregonfrogs)
```


Let's see how to use it
```{r}
oregonfrogs_raw %>%
  dplyr::select(SurveyDate, Frequency, UTME_83, UTMN_83) %>%
  head()
```

How to use `longlat_to_utm()` to transform the UTM coordinates to LongLat:

```{r}
oregonfrogs_raw %>%
  dplyr::select(SurveyDate, Frequency, UTME_83, UTMN_83) %>%
  utm_to_longlat(utm_crs = "+proj=utm +zone=10",
                 longlat_crs = "+proj=longlat +datum=WGS84") %>%
  head()
```

Here a simple usage of the spatial coordinates is done for showing possible patterns in frogs movements. The `ggplot2::geom_smooth()` function shows two models the linear model (`lm`) and the LOESS.
```{r}
library(ggplot2)
oregonfrogs_raw %>%
  dplyr::select(SurveyDate, Frequency, UTME_83, UTMN_83) %>%
  utm_to_longlat(utm_crs = "+proj=utm +zone=10",
                 longlat_crs = "+proj=longlat +datum=WGS84") %>%
  ggplot(aes(x = X, y = Y)) +
  geom_point(
    alpha = 0.5,
    shape = 21,
    stroke = 0.5,
    fill = "#32cd32"
  ) +
  geom_smooth(color = "#cf5f25", alpha = 0.2) +
  geom_smooth(method = "lm",
              color = "#a8bd3a",
              alpha = 0.2) +
  labs(x = "Longitude", y = "Latitude") +
  theme_bw()
```

